On the validity of the Born approximation for Weak Lensing observables

### 2016-10-26 Paper/paper.tex
use seaborn for plotting
first results

\begin{equation}
\kappa = \kappa^{\rm born} + \kappa^{\rm lens-lens} + \kappa^{\rm geodesic} + O(\Phi^3)
\end{equation}
%
The second order corrections can be expressed as double integrals along the unperturbed line of sight

@@

\begin{equation}
\label{sim:ll}
\kappa^{\rm lens-lens}(\pt,\chi_s) = \int_0^{\chi_s}d\chi\int_0^\chi d\chi' \chi\chi' W(\chi,\chi_s)W(\chi',\chi_s)\Phi_{ij}(\chi\pt,\chi)\Phi_{ij}(\chi'\pt,\chi')
\end{equation}

\begin{equation}
\label{sim:gp}
\kappa^{\rm geodesic}(\pt,\chi_s) = \frac{3H_0^2\Omega_m}{2c^2}\int_0^{\chi_s}d\chi\int_0^\chi d\chi' \frac{\chi\chi'}{a(\chi')} W(\chi,\chi_s)W(\chi',\chi_s)\nabla_\perp\Phi(\chi\pt,\chi)\cdot\nabla_\perp\delta(\chi'\pt,\chi')
\end{equation}

@@

\begin{equation}
\label{par:linapprox}
\bb{d}(\bb{p}) = \bb{d}_0 + \bb{M}(\bb{p}-\bb{p}_0)
\end{equation}
%
the maximum likelihood parameter estimator (\ref{par:fisherest}) is given by (see \citep{DodelsonSchneider13,PetriVariance} for example)
\begin{equation}
\label{par:fisherest}
\bbh{p} = \bb{p}_0 + (\bb{M}^T\bb{\Psi M})^{-1} \bb{M}^T\bb{\Psi} (\bbh{d}_{\rm obs}-\bb{d}_0)
\end{equation}
%
where $\bb{\Psi}=(\bb{C}^{\bb{d}\bb{d}})^{-1}$. The linearity assumption (\ref{par:linapprox}) is justified when the scatter of the measure feature $\bbh{d}_{\rm obs}$ around the expansion point $\bb{d}_0$ is expected to be small. This is true for upcoming large area surveys such as LSST \citep{LSST}, WFIRST \citep{WFIRST} and Euclid \citep{Euclid}. In this work we use $N$--body simulations coupled with the \ttt{LensTools} routines to simulate multiple realizations of a convergence field $\h{\kappa}(\pt)$ for different combinations of the cosmological parameter triplet $(\Omega_m,w,\sigma_8)$. We choose the values $\bbh{p}_0=(0.26,-1,0.8)$ as our fiducial cosmological parameters and vary one parameter at a time as in Table \ref{tab:cosmopar} to measure the feature derivatives $\bb{M}$.

@@

\begin{tabular}{c|c|c|c}
$\Omega_m$ & $w$ & $\sigma_8$ & $\kappa$ realizations \\ \hline \hline
\multicolumn{4}{c}{\textbf{Fiducial}} \\ \hline
0.26 & -1 & 0.8 & 8192 \\ \hline

\multicolumn{4}{c}{\textbf{Variation 1}} \\ \hline
0.29 & -1 & 0.8 & 1024 \\
0.26 & -0.8 & 0.8 & 1024 \\
0.26 & -1 & 0.9 & 1024 \\ \hline

\multicolumn{4}{c}{\textbf{Variation 2}} \\ \hline
0.23 & -1 & 0.8 & 1024 \\
0.26 & -1.2 & 0.8 & 1024 \\
0.26 & -1 & 0.7 & 1024 \\ \hline

@@

\begin{equation}
\label{par:moments}
\begin{matrix}
\pmb{\mu} = (\pmb{\mu}_2,\pmb{\mu}_3,\pmb{\mu}_4) \\ \\
\pmb{\mu}_2 = \left(\langle\kappa^2\rangle,\langle\vert\nabla\kappa\vert^2\rangle\right) \\ \\
\pmb{\mu}_3 = \left(\langle\kappa^3\rangle,\langle\kappa^2\nabla^2\kappa\rangle,\langle\kappa\vert\nabla\kappa\vert^2\rangle\right) \\ \\
\pmb{\mu}_4 = \left(\langle\kappa^4\rangle_c,\langle\kappa^3\nabla^2\kappa\rangle_c,\langle\kappa^2\vert\nabla\kappa\vert^2\rangle_c,\langle\vert\nabla\kappa\vert^4\rangle_c\right) \\
\end{matrix}
\end{equation} 
%
Where the expectation values are taken with respect to pixels and the subscript $c$ indicated that we are considering the connected components of the quartic moments $\pmb{\mu}_4$. In the next section we outline our main results.

@@

\section{Results}
%

@@

@@

>>>>

\begin{equation}
\label{sim:pb2}
\kappa = \kappa^{\rm born} + \kappa^{\rm ll} + \kappa^{\rm geo} + O(\Phi^3)
\end{equation}
%
The second order corrections can be expressed as double integrals along the unperturbed line of sight

@@

\begin{equation}
\label{sim:ll}
\kappa^{\rm ll}(\pt,\chi_s) = \int_0^{\chi_s}d\chi\int_0^\chi d\chi' \chi\chi' W(\chi,\chi_s)W(\chi',\chi_s)\Phi_{ij}(\chi\pt,\chi)\Phi_{ij}(\chi'\pt,\chi')
\end{equation}

\begin{equation}
\label{sim:gp}
\kappa^{\rm geo}(\pt,\chi_s) = \frac{3H_0^2\Omega_m}{2c^2}\int_0^{\chi_s}d\chi\int_0^\chi d\chi' \frac{\chi\chi'}{a(\chi')} W(\chi,\chi_s)W(\chi',\chi_s)\nabla_\perp\Phi(\chi\pt,\chi)\cdot\nabla_\perp\delta(\chi'\pt,\chi')
\end{equation}

@@

\begin{equation}
\label{par:linapprox}
\bb{d}(\bb{p}) = \bb{d}_0 + \bb{M}(\bb{p}-\bb{p}_0)
\end{equation}
%
the maximum likelihood parameter estimator (\ref{par:fisherest}) is given by (see \citep{DodelsonSchneider13,PetriVariance} for example)
\begin{equation}
\label{par:fisherest}
\bbh{p} = \bb{p}_0 + (\bb{M}^T\bb{\Psi M})^{-1} \bb{M}^T\bb{\Psi} (\bbh{d}_{\rm obs}-\bb{d}_0)
\end{equation}
%
where $\bb{\Psi}=(\bb{C}^{\bb{d}\bb{d}})^{-1}$. The linearity assumption (\ref{par:linapprox}) is justified when the scatter of the measure feature $\bbh{d}_{\rm obs}$ around the expansion point $\bb{d}_0$ is expected to be small. This is true for upcoming large area surveys such as LSST \citep{LSST}, WFIRST \citep{WFIRST} and Euclid \citep{Euclid}. In this work we use $N$--body simulations coupled with the \ttt{LensTools} routines to simulate multiple realizations of a convergence field $\h{\kappa}(\pt)$ for different combinations of the cosmological parameter triplet $(\Omega_m,w_0,\sigma_8)$. We choose the values $\bbh{p}_0=(0.26,-1,0.8)$ as our fiducial cosmological parameters and vary one parameter at a time as in Table \ref{tab:cosmopar} to measure the feature derivatives $\bb{M}$.

@@

\begin{tabular}{c|c|c|c}

$\Omega_m$ & $w_0$ & $\sigma_8$ & $\kappa$ realizations \\ \hline \hline
\multicolumn{4}{c}{\textbf{Fiducial}} \\ \hline
0.26 & -1 & 0.8 & 8192 \\ \hline

\multicolumn{4}{c}{\textbf{Variations}} \\ \hline
0.29 & -1 & 0.8 & 1024 \\
0.26 & -0.8 & 0.8 & 1024 \\
0.26 & -1 & 0.9 & 1024 \\ \hline

@@

\begin{equation}
\label{par:moments}
\begin{matrix}
\pmb{\mu} = (\pmb{\mu}_2,\pmb{\mu}_3,\pmb{\mu}_4) \\ \\
\pmb{\mu}_2 = \left(\langle\kappa^2\rangle,\langle\vert\nabla\kappa\vert^2\rangle\right) \\ \\
\pmb{\mu}_3 = \left(\langle\kappa^3\rangle,\langle\kappa^2\nabla^2\kappa\rangle,\langle\vert\nabla\kappa\vert^2\nabla^2\kappa\rangle\right) \\ \\
\pmb{\mu}_4 = \left(\langle\kappa^4\rangle_c,\langle\kappa^3\nabla^2\kappa\rangle_c,\langle\kappa\vert\nabla\kappa\vert^2\nabla^2\kappa\rangle_c,\langle\vert\nabla\kappa\vert^4\rangle_c\right) \\
\end{matrix}
\end{equation} 
%
Where the expectation values are taken with respect to pixels and the subscript $c$ indicated that we are considering the connected components of the quartic moments $\pmb{\mu}_4$. In the next section we outline our main results.

@@

\section{Results}
% 
In this section we presents the results of this work, regarding the WL feature forward model accuracy and the cosmological parameter bias induced by the Born approximation.

@@

In this paragraph we compare the accuracy of the Born approximation in predicting WL image features, in particular we focus on the $\kappa$ power spectrum and the 9 moments described in equation (\ref{par:moments}). Figure \ref{fig:psResiduals} shows the auto power spectra of $\kappa^{\rm born},\kappa^{\rm lens-lens},\kappa^{\rm geodesic}$ as defined in equation (\ref{sim:pb2}). At lowest order the lensing potential contributes to the $\kappa$ power spectrum quadratically. The first post--Born correction to $P^{\kappa\kappa}$ comes to $O(\Phi^3)$ and is given by
\begin{equation}
\label{res:powercorr}
P^{\kappa\kappa} = P^{\kappa\kappa}_{\rm born,born} + 2P^{\kappa\kappa}_{\rm born,ll} + 2P^{\kappa\kappa}_{\rm born,geo} + O(\Phi^4)
\end{equation}
%
Where we defined the $\kappa$ cross spectra as
\begin{equation}
\label{res:powercross}
\langle\tilde{\kappa}^{a}(\pmb{\ell})\tilde{\kappa}^{b,*}(\pmb{\ell}')\rangle = (2\pi)^2\delta^D(\pmb{\ell}+\pmb{\ell}')P_{a,b}^{\kappa\kappa}(\ell)
\end{equation} 
%
Figure \ref{fig:psResiduals} shows the magnitudes of the $O(\Phi^3)$ post--Born corrections to the convergence power spectrum.

\begin{figure*}
\begin{center}
\includegraphics[scale=0.3]{Figures/powerResiduals.eps}
\end{center}
\caption{Power spectrum residuals}
\label{fig:psResiduals}
\end{figure*}

\begin{figure*}
\begin{center}
\includegraphics[scale=0.25]{Figures/pdfMoments.eps}
\end{center}
\caption{PDF of $\kappa$ moments}
\label{fig:pdfMoments}
\end{figure*}

@@

\begin{figure*}
\begin{center}
\includegraphics[scale=0.25]{Figures/bias_convergence_power_s0_nb100.eps}
\includegraphics[scale=0.25]{Figures/bias_convergence_moments_s0_nb9.eps}
\includegraphics[scale=0.25]{Figures/bias_convergence_momentsSN_s0_nb9.eps}
\end{center}
\caption{Parameter bias}
\label{fig:parbias}
\end{figure*}
